---
title: "Examining results Stochastic MT3DMS run"
author: "G. Bier"
output:
  html_document:
    toc: yes
    df_print: paged
  html_notebook:
    toc: yes
---

# Introduction
With this markdown you can visualize the different central mass moments from the stochastic MT3DMS model run.

With the `COVxy` you can run an animation showing how the plume is distributed in the domain during time.


# Read the data
The `read_stoch_MT3DMS.R` script should be run first to determine the central mass moments for all the stochastic MT3DMS
The data is stored in a csv file with the name is based on number of columns, rows, mean_k,sd_k,l_y and the number of simulations like e.g.  `144_72_k70_sdk3_ly10_nrsim250_MT3DMS.csv`. It will also produces a file with the dimensions of the model, e.g. `144_72_k70_sdk3_ly10_nrsim250_MT3DMS.dim`.

```{r}
rm(list = ls())  # clean the memory
library(gifski) #package required for animating the displacement of the plumes
cmm_file = "144_72_k70_sdk3_ly10_nrsim250_MT3DMS.csv" #name of the file containing the central mass moments
mm_data = read.csv(file = cmm_file)  #read the data coming from the stochastic run 
####dim_file and dim.dat are only here for proper scaling of the plume shapes
dim_file = paste0(substr(cmm_file,1,nchar(cmm_file) -3),"dim")
dim.dat = read.table(file = dim_file)
```

# Calculate some statistics of the time series
It could be interesting to have a look at the average and standard deviation central mass moments

```{r results='hide'}
# setting up the containers for the mean and standard deviation of the (central) mass moments
M0 = c()
M0mean = c()
M0sd = c()
M1x = c()
M1xmean = c()
M1xsd = c()
M1y = c()
M1ymean = c()
M1ysd = c()
M2x = c()
M2xmean = c()
M2xsd = c()
M2y = c()
M2ymean = c()
M2ysd = c()

##create indexes for the different central mass moments
M0index = grep("M0",names(mm_data))
M1xindex = grep("M1x",names(mm_data))
M1yindex = grep("M1y",names(mm_data))
M2xindex = grep("M2x",names(mm_data))
M2yindex = grep("M2y",names(mm_data))
##
## now calculate the mean and standard deviation of the central mass moments
for (i in 1:length(mm_data[,1])) {
  print(paste("reading time step :",i))
  M0 = as.matrix(mm_data[i,M0index])
  M1x = as.matrix(mm_data[i,M1xindex])
  M1y = as.matrix(mm_data[i,M1yindex])
  M2x = as.matrix(mm_data[i,M2xindex])
  M2y = as.matrix(mm_data[i,M2yindex])
  M0mean = c(M0mean,mean(M0))
  M0sd = c(M0sd,sd(M0))
  M1xmean = c(M1xmean,mean(M1x))
  M1xsd = c(M1xsd,sd(M1x))
  M1ymean = c(M1ymean,mean(M1y))
  M1ysd = c(M1ysd,sd(M1y))
  M2xmean = c(M2xmean,mean(M2x))
  M2xsd = c(M2xsd,sd(M2x))
  M2ymean = c(M2ymean,mean(M2y))
  M2ysd = c(M2ysd,sd(M2y))

}
```

## Adding two plot functions for evaluation

`plotMM_mean_sd(mm = string)` plots the mean and standard deviations of that mean of the chosen central mass moment
The argument `mm` should be a string holding the name of this central mass moment variable.

`plotMM_all(mm = string)` plot all the time series of `mm`
The argument `mm` should be a string holding the name of this central mass moment variable.


```{r}
plotMM_mean_sd = function(mm)
{
  if(!is.character(mm))
  {
    stop("Variable should be a string, so with quotes")
  }
  MMmean= (paste0(mm,"mean"))
  mean = eval(as.name(MMmean))
  MMsd = paste0(mm,"sd")
  sd = eval(as.name(MMsd))
  if (mm == "M2x" | mm == "M2y"){
    ylabel = paste(mm, " in Length^2 (m2)")} 
  if (mm == "M0"){
    ylabel = paste(mm, "in Mass (kg)")}
  if(mm == "M1x" | mm == "M1y"){
    ylabel = paste(mm, "in Length (m)")}
  
  
  plt.range = c(min(mean)-max(sd),max(mean)+max(sd))
 plot(mm_data$pts,mean, col = "blue",lwd =2,type = "o",
      ylim = plt.range, ylab = ylabel, xlab = "Time in d",
      main = paste("Average and stdev of (central) MM: ",mm))
 lines(mm_data$pts,mean+sd, col = "red",lty = "dashed")
 lines(mm_data$pts,mean-sd, col = "red",lty = "dashed")
 grid()
}

plotMM_mean_sd("M0")
plotMM_mean_sd("M1x")
plotMM_mean_sd("M1y")
plotMM_mean_sd("M2x")
plotMM_mean_sd("M2y")


nr.set = (length(mm_data[1,]) - 2)/6 
col.set = rainbow(n=nr.set)

###just for checking the colors
plot(1,1,type = "p", xlim = c(0,nr.set),ylim = c(0,nr.set),
     pch = 16, cex=2,col=col.set[1])
for (i in 2:length(col.set))
{
points(i,i,pch = 16, cex=2,col=col.set[i] )
}
###just for checking the colors

plotMM_all = function(mm)
{
  if(!is.character(mm))
  {
    stop("Variable should be a string, so with quotes")
  }
  
  if (mm == "M2x" | mm == "M2y"){
    ylabel = paste(mm, " in Length^2 (m2)")} 
  if (mm == "M0"){
    ylabel = paste(mm, "in Mass (kg)")}
  if(mm == "M1x" | mm == "M1y"){
    ylabel = paste(mm, "in Length (m)")}

  
  
 mm.index = grep(pattern = mm, x=names(mm_data))
 plot(mm_data$pts,mm_data[,mm.index[1]],
      ylim = c(min(mm_data[,mm.index]),max(mm_data[,mm.index])),
      ylab = ylabel, xlab = "Time in d",
      col = col.set[1],type="l")
 title(paste("Central mass moments of ", mm))
 for (i in 2:length(mm.index))
 {
   lines(mm_data$pts,mm_data[,mm.index[i]],col=col.set[i])
 }
   grid()
}

plotMM_all("M0")
plotMM_all("M1x")
plotMM_all("M1y")
plotMM_all("M2x")
plotMM_all("M2y")
```
## Addtional plot for **D11** macroscopic dispersion

```{r}
dM2xmean = diff(M2xmean)
dpts = diff(mm_data$pts)
plot(mm_data$pts[-length(dpts)], dM2xmean/(2*dpts), type = "o",lwd = 2)
xdata = mm_data$pts[-length(dpts)]
ydata = diff(M2xmean)/(2*dpts)
fitD11_lin = lm(ydata ~ xdata)
fitD11_ln = lm(ydata ~log(xdata))
abline(fitD11_lin, col = "blue")
abline(fitD11_ln, col = "red")
lines(fitD11_ln$fitted.values,col = "red")
grid()

sdM2xmean = sqrt(M2xmean)
plot(mm_data$pts,sdM2xmean, type = "o",col = "blue",lwd = 2,
     main = "Standard deviation of M2x, spreading in x in m")
grid()
fit_sdMean = lm(sdM2xmean ~ mm_data$pts)
abline(fit_sdMean, col = "red")

plot(sqrt(mm_data$pts),sdM2xmean, type = "o",col = "blue",lwd = 2,
     main = "Standard deviation of M2x, spreading in x in m")
grid()
fit_sdMean = lm(sdM2xmean ~ sqrt(mm_data$pts))
abline(fit_sdMean, col = "red")

```



## Animation of the plume displacement

Below an animation of the displace of the "statistical" plume which is based on the correlation matrix
This matrix is calculated by

$$ 
COR(x,y) = \frac{COV(x,y)}{\sigma x \sigma y}\\
 \mathbf C_{x,y} = \begin{bmatrix}
 1 & COR(x,y)\\
 COR(x,y) & 1\\
 \end {bmatrix}
$$
To create the "statistical plume" the packages `ellipse` need to be installed and loaded. The `gifski` package is also required  for creating the animation.
The animation within the html document need to knitted (cannot be generated with *preview* only)

```{r animation.hook='gifski', interval= 0.25}

#their should be a *.dim file available containing the dimensions of the  model


library(ellipse)
cov.index = grep("Covxy", names(mm_data))
m2x.index = grep("M2x", names(mm_data))
m2y.index = grep("M2y", names(mm_data))
m1x.index = grep("M1x",names(mm_data))
m1y.index = grep("M1y", names(mm_data))

col.set = rainbow(nr.set, alpha = 0.2)
col.center = rainbow(nr.set)
# col.set = topo.colors(nr.set,alpha =0.2)
# col.center = topo.colors(nr.set,alpha = 1)
# col.set = hcl.colors(n = nr.set,palette = "viridis",alpha =0.2)
# col.center = hcl.colors(n = nr.set,palette = "viridis")
# col.set = cm.colors(nr.set,alpha = 0.2)
# col.center = cm.colors(nr.set)
# col.set = heat.colors(nr.set, alpha = 0.2)
# col.center = heat.colors(nr.set)

old.par = par(no.readonly = TRUE)  #to save the original plotting parameters

y.model = dim.dat$nr.rows*dim.dat$dy+dim.dat$dy/2
x.model = dim.dat$nr.cols*dim.dat$dx+dim.dat$dx/2

for (i in 1:length(mm_data$pts))#requires gifski and ellipse package!!
{ 
  plot(1,1, type = "n", ylim=c(0,y.model),xlim=c(0,x.model),xlab= "distance (m)",ylab = "height (m)")
 
  for (p in 1: length(cov.index))
  {
    M1x = mm_data[,m1x.index[p]]
    M1y = mm_data[,m1y.index[p]]
    M2x = mm_data[,m2x.index[p]]
    M2y = mm_data[,m2y.index[p]]
    COVxy = mm_data[,cov.index[p]]
    
    
    corxy = COVxy[i]/sqrt(M2x[i] * M2y[i])
    M  = matrix(c(1,corxy,corxy,1),ncol = 2, byrow=TRUE) #the correlation matrix seems to be required for ellipse
    par(new = TRUE)
    #plot(M1x[1:i],M1y[1:i],type="l",xlim=c(2,40),asp=1,col=col.set[p],lwd=2,
    #     main=paste("time=",mm_data$pts[i]),xlab="x",ylab="y")
    title(main = paste("time=",format(mm_data$pts[i], digits = 2)))
    lines(M1x[1:i],M1y[1:i],col=col.set[p])
    points(M1x[i],M1y[i],pch=20,col=col.center[p],cex=1)
    #polygon(ellipse(M,scale=c(M2x[i],M2y[i]),centre=c(M1x[i],M1y[i]),level=0.68,fill=TRUE),col=rgb(0,0,1,0.2),border=NA)
    #scale.ellipse = sqrt(M2x[i]+M2y[i])
    #polygon(ellipse(M,scale=c(M2x[i]/M2xmean[i],M2y[i]/M2ymean[i]),centre=c(M1x[i],M1y[i]),level=0.68,fill=F),col=col.set[p],border=NA)
    polygon(ellipse(M,scale=c(sqrt(M2x[i]),sqrt(M2y[i])),centre=c(M1x[i],M1y[i]),level=0.68,fill=T),col=col.set[p],border=NA)
    grid(col="black")
  }
}
 par = old.par

```
### plotting individual plumes

```{r}

# COVxy = mm_data[,cov.index[3]]
# for (i in 1:length(mm_data$pts))#requires gifski and ellipse packages!!
# {
#   corxy = COVxy[i]/sqrt(M2x[i] * M2y[i])
#   print(corxy)
#   M  = matrix(c(1,corxy,corxy,1),ncol = 2, byrow=TRUE) #the correlation matrix seems to be required for ellipse
# 
#   plot(M1x[1:i],M1y[1:i],type="l",xlim=c(2,40),asp=1,col=col.set[1],lwd=2,
#        main=paste("time=",mm_data$pts[i]),xlab="x",ylab="y")
#   points(M1x[i],M1y[i],pch=20,col=col.set[1],cex=2)
#   #polygon(ellipse(M,scale=c(M2x[i],M2y[i]),centre=c(M1x[i],M1y[i]),level=0.68,fill=TRUE),col=rgb(0,0,1,0.2),border=NA)
#   polygon(ellipse(M,scale=0.1*c(M2x[i],M2y[i]),centre=c(M1x[i],M1y[i]),level=0.68,fill=TRUE),col=col.set[1],border=NA)
#   grid(col="black")
# }

```


