---
title: "Spatially correlated random fields"
output: html_notebook
---

# 1. Introduction

Based on url: <http://santiago.begueria.es/2010/10/generating-spatially-correlated-random-fields-with-r/>


**Unconditional Gaussian simulation using gstat**

Spatial modellers commonly use the term unconditional Gaussian simulation to refer to the process of generating spatially correlated random fields. For comparison purposes, the Figure 0 shows an example of a random field with no spatial correlation. The value of the measured property at one given cell is completely independent of the values of that property at neighbouring cells. This situation is very seldom (or I perhaps should better say never) found in nature, since spatially distributed variables always have a certain level of autocorrelation, i.e. co-variance between neighbours. This is often mentioned as the 'geographical law', stating that closer locations tend to have similar properties.

Generating spatially correlated random fields is interesting because it makes it possible testing different issues related to the statistical analysis of spatial data.



## 1.1. Generating the spatial field

We are going to use the `gstat` library for generating the field and the `sp` library for some manipulations with spatial data (e.g. `gridded, spplot` ), so we start by loading them:

```{r}
rm(list=ls()) #cleaning the environment
library(gstat)#load packages
library(sp)
```

### 1.1.1. Set the seed

With the function below you can set the seed (a random number) so you can can reproduce exactly the same k-fields

```{r}
set.seed(01022024)
```


We create a `num_col` x `num_row` grid, and we convert it into a data frame (xyz structure) by taking all possible combinations of the x and y coordinates:

```{r}

num_row = 72
num_col = 144
num_values = num_row*num_col
  
xy <- expand.grid(1:num_col, 1:num_row) #to set up the grid 
names(xy) <- c("x","y")
```

 

### 1.1.1. Spatial properties random field

The statistical properties are given below:

-   `k_mean` : the mean of the random field (k for hydraulic conductivity)

-   `k_sd` : the standard deviation around the mean k value

The real mean and standard deviation need to be scaled since the `ln(k)` is a normal distribution resulting in a lognormal `k` distribution to be used in MODFLOW.

#### 1.1.1.1. Set the mean and standard deviation of the conductivity field

```{r}
# statistics
k_mean = 7.50 
k_sd = 3.0

scaled_mean = log(k_mean^2/sqrt(k_mean^2+k_sd^2))
scaled_stdev = sqrt(log((k_mean^2+k_sd^2)/k_mean^2))
#scaled_stdev = sqrt(log(1 + (k_mean^2+k_sd^2)/k_mean^2)) #isn't this the correct transformation?


```

### 1.1.2. Simulation model

Defining the spatial model as a `gstat` object (text from reference):

`g.dummy <- gstat(formula=z~1, locations=~x+y, dummy=T, beta=1, model=vgm(psill=0.025, range=5, model='Exp'), nmax=20)`

with:

-   `g.dummy` : the object which contains the info for the simulation of the spatially correlated random field

-   `formula` :

    defines the dependent variable (z) as a linear model of independent variables. For ordinary and simple kriging we can use the formula `z~1`; for simple kriging it is necessary to define a beta parameter too (see below); for universal kriging, if `z` is linearly dependent on `x` and `y` use the formula `z~x+y`. We are using simple kriging here.

-   `locations` define the data coordinates, e.g. `~x+y` in our case here.

-   `dummy` is a logical value, and it needs to be TRUE for unconditional simulation.

-   `beta` is used only for simple kriging, and is a vector with the trend coefficients (including an intercept); if no independent variables are defined the model only contains an intercept, i.e. the simple kriging mean.

-   `model` defines the variogram model, as defined by a call to `vgm`.

    -   `psill` : (scaled) variance of the k field (after this, values are uncorrelated). Also called contribution

    -   `range` : correlation length alike (distance at which samples/values become independent of each other)

    -   `model` : variogram model type; `Exp` :Exponential, `Gau` : Gaussian, `Sph` : Spherical and more

-   `nmax` defines the number of nearest observations that should be used for a kriging prediction or simulation.

#### 1.1.2.1. Setting the autocorrelation length

```{r}

range = 5 #low value means hardly any correlation length
nmax = 5 # low values with low range means very random simulation

```

creation of the model

```{r}
g.dummy <- gstat(formula=z~1, locations=~x+y, dummy=T, beta=scaled_mean, model=vgm(psill=scaled_stdev^2, model="Exp",range=range), nmax=nmax)
```
 
 

2.  Running Simulations

With `g.dummy` the model is set up to start simulating.

-   `nsim` : the number of simulations/fields to generate

-   `predict` : function to simulate/generate the field based on the chosen kriging method

    -   `g.dummy` : the model object we have set up

    -   `newdata` : data frame created earlier to write the results to (so a grid like data frame)

-   `gridded` :

```{r}
nsim = 50 # number of realizations
k_fields <- predict(g.dummy, newdata=xy, nsim = nsim)
gridded(k_fields) = ~x+y


```

 

3.  Viewing and analyzing the results

To see if the results are as expected it is practical to have a look at the results with use of e.g. histograms of the data.

```{r}
#here we show the histograms of the data
spdf <- as(k_fields, "data.frame")
for (col in names(spdf)[-1]) {  # assuming the first column is spatial data
  hist(spdf[[col]], main = paste("Histogram of ln(k)", col), xlab = col, ylab = "Frequency", col = "lightblue", border = "black")
}
for (col in names(spdf)[-1]) {  # assuming the first column is spatial data
  hist(exp(spdf[[col]]), main = paste("Histogram of k", col), xlab = col, ylab = "Frequency", col = "blue", border = "black")
}
#here we show the fields in ln(k) and k space
spplot(k_fields) #plotting the fields in ln space
spplot(k_fields,formula= exp(z)~x+y|name) #plotting the fields in k space
```

# 4. Exporting the data to a txt file

The data contained in the k_fields structure will be exported to a simple one column per simulation format ASCII text file.

```{r}
data_kvalue = as.data.frame(matrix(ncol=nsim, nrow=num_values))
for (i in 1:nsim){
  data_kvalue[i] = round(exp(k_fields@data[,i]),2) #setting up the data.frame for     the data
}
#writing the data fo a file
write.table(data_kvalue,file = paste0("k_col",num_col,"_row",num_row,"k",k_mean,"sd_k",k_sd,"Ly",range,"simnr",nsim,".txt"), 
            append = FALSE, sep = "\t", dec = ".", row.names= FALSE, col.names=TRUE)
```

 
